Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar

### 2019-01-29 accessPEP.Rmd
reorganizing for meeting Jan 30th

@@

### GIS Data
We use the global friction surface for 2015 generated by the Malaria Atlas Project ( https://map.ox.ac.uk/research-project/accessibility_to_cities/, Weiss et al. 2015,) and GPS points of clinics to estimate the travel time to the nearest ARMC for the country of Madagascar at a 1 x 1 km scale. We then calculated a weighted average of travel times by human population to the commune level, using administrative shapefiles available trhough the UN Office for the Coordination of Humanitarian Affairs. For each clinic, we defined the catchment area as all districts for which the clinic was the closest ARMC. Human population estimates were taken from the 2015 UN adjusted population projections from World Pop (www.worldpop.org, Linaird et al. 2012) and also aggreagated to the commune level
<br>

### Model sensitivity/validation
- Out-of-fit check for Tana + Fianarantsoa?
- Compare to null models 
- Look at sensitivity to inclusion criteria of days and exclusion of contacts

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
rm(list = ls())

@@

@@

We attempted to exclude these by looking at the distribution of patients reporting per day to a given clini, excluding patients reporting on any dates that had greater than 3 standard deviations above the mean number of patients reporting per day (Figure 3A). We validated this using the Moramanga ARMC data for which we had details on the type of exposure, and found that setting the threshold to 3 standard deviations resulted in approximately 50% of known contacts excluded, with only 2% of non-contacts excluded (Figure 3B). For the national data for which a subset of patient forms were noted to be contacts, we found that our exclusion criteria of 3 SDs identified `r round((ctar_check %$% length(exclude_bycomment[exclude_bycomment == 1 & exclude_bydate == 1]))/length(ctar_check %$% exclude_bycomment[exclude_bycomment == 1])*100, 2)` of known contacts. We further excluded known contacts which were not identified based on the daily distribution of patients, resulting in the exclusion of approximately `r round((ctar_check %$% length(exclude_bycomment[exclude_bycomment == 1 | exclude_bydate == 1]))/nrow(ctar_check)*100, 2)`% of patient records from the national data.

@@

ggplot(exps, aes(x = reorder(dist_name, ttimes_weighted_dist), y = bites/pop*1e5, color = district)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  scale_color_manual(values = col_vals) + 
  ylab("Annual bites per 100k") +
  xlab("") +
  coord_flip()

>>>>

```{r set up, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
rm(list = ls())

@@

@@

### GIS Data
We use the global friction surface for 2015 generated by the Malaria Atlas Project ( https://map.ox.ac.uk/research-project/accessibility_to_cities/, Weiss et al. 2015,) and GPS points of clinics to estimate the travel time to the nearest ARMC for the country of Madagascar at a 1 x 1 km scale. We then calculated a weighted average of travel times by human population to the commune level, using administrative shapefiles available trhough the UN Office for the Coordination of Humanitarian Affairs. For each clinic, we defined the catchment area as all districts for which the clinic was the closest ARMC. Human population estimates were taken from the 2015 UN adjusted population projections from World Pop (www.worldpop.org, Linaird et al. 2012) and also aggreagated to the commune level

@@

We attempted to exclude these by looking at the distribution of patients reporting per day to a given clinic, excluding patients reporting on any dates that had greater than 3 standard deviations above the mean number of patients reporting per day (Figure 3A). We validated this using the Moramanga ARMC data for which we had details on the type of exposure, and found that setting the threshold to 3 standard deviations resulted in approximately 50% of known contacts excluded, with only 2% of non-contacts excluded (Figure 3B). For the national data for which a subset of patient forms were noted to be contacts, we found that our exclusion criteria of 3 SDs identified `r round((ctar_check %$% length(exclude_bycomment[exclude_bycomment == 1 & exclude_bydate == 1]))/length(ctar_check %$% exclude_bycomment[exclude_bycomment == 1])*100, 2)` of known contacts. We further excluded known contacts which were not identified based on the daily distribution of patients, resulting in the exclusion of approximately `r round((ctar_check %$% length(exclude_bycomment[exclude_bycomment == 1 | exclude_bydate == 1]))/nrow(ctar_check)*100, 2)`% of patient records from the national data.

@@

ggplot(exps, aes(x = reorder(dist_name, ttimes_weighted_dist), y = bites/pop*1e5, color = district)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  scale_color_manual(values = col_vals) + 
  ylab("Annual bites per 100k") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
